Using Google Community Mobility Reports to investigate the incidence of COVID-19 in the United States

### 2020-05-19 Covid-19-Google-CMR-US/Response-to-Reviewers-v3.Rmd

```{r model-results-pct-increase, echo = FALSE}
data.frame(Variable = c(#"date",
                        #"date^2",
                        #"parks^2",
                        "parks",
                        "parks x work",
                        "work",
                        "work^2",
                        "parks x date",
                        "parks x date^2",
                        "work x date^2",
                        "NY",
                        "NY x date"),
           Coefficient = as.numeric(coefficients(mod_pct_increase)),
           pval = as.numeric(summary(mod_pct_increase)$coefficients[,4])) %>%
  mutate(pval = ifelse(pval >= 0.001, round(pval, 4), "<0.001")) %>%
  kable("latex",
        booktabs = TRUE,
        digits = 4,
        caption = "\\label{tab:model-results-pct-increase}Dependent variable is daily percent increase in total cases.",
        align = c("l", "c", "c"),
        col.names = c("Variable", "Coefficient Estimate", "p-value")) %>%
  kable_styling(latex_options = c("striped"), font_size = 7) %>%
  footnote(general = c(paste("Coefficient of Determination $R^2$=", round(summary(mod_pct_increase)$r.squared, 3)),
                       paste("Adjusted Coefficient of Determination $R^2$=", round(summary(mod_pct_increase)$adj.r.squared, 3)),
                       paste("Standard Error $\\\\sigma$=", round(summary(mod_pct_increase)$sigma, 3))),
           escape = FALSE)
```

>\textcolor{blue}{The examples above show that using the day-to-day percent daily change in new cases and percent daily increase in total cases does not result in better models. The point about controlling for population still stands. This made me think of a different dependent variable, namely incidence per 100,000:}

$$
incidence_{i,t} = \frac{Cases_{i,t}}{population_{i}/100,000}
$$

>\textcolor{blue}{Incidence normalizes the number of cases and accounts for the size of the population at risk. This variable is log-transformed to scale it (it has a long tail, as you would expect). The correlations with the mobility indicators are shown in Table \ref{tab:correlation-analysis-log-incidence}. Using the log of incidence as the dependent variable results in a somewhat more parsimonious model with better goodness of fit (see Table \ref{tab:model-results-log-incidence}). Based on these results, this is the model that I report in the paper now. It is worthwhile noting that qualitatively the results are very similar, but with a slightly smaller model that also performs better.}

```{r check-correlations-log-incidence, echo=FALSE}
#Check the correlation between the mobility indicators and log of incidence:
c19 %>% 
  select(log_incidence, retail, groceries, parks, transit, work, residential) %>%
  cor(use = "pairwise.complete.obs") %>%
  kable("latex",
        booktabs = TRUE,
        digits = 2,
        caption = "\\label{tab:correlation-analysis-log-incidence}Simple correlation between log(incidence) and the mobility indicators")  %>%
  kable_styling(latex_options = c("striped"), font_size = 7) %>%
  footnote(general = "All mobility indicators are lagged 11-day moving averages")
```

```{r model-estimation-log-incidence, include=FALSE}
mod1 <- lm(log_incidence ~  0 +
                         day +
                         day2 +
                         parks2 + 
                         parks + 
                         parksXwork +
                         work +
                         work2 +
                         parksXday + 
                         parksXday2 +
                         #workXday +
                         workXday2 +
                         NY, #+ 
                         #NYxday, 
                       data = c19)
summ(mod1)
```

```{r model-results-log-incidence, echo = FALSE}
data.frame(Variable = c("date",
                        "date^2",
                        "parks^2",
                        "parks",
                        "parks x work",
                        "work",
                        "work^2",
                        "parks x date",
                        "parks x date^2",
                        "work x date^2",
                        "NY"),
           Coefficient = as.numeric(coefficients(mod1)),
           pval = as.numeric(summary(mod1)$coefficients[,4])) %>%
  mutate(pval = ifelse(pval >= 0.001, round(pval, 4), "<0.001")) %>%
  kable("latex",
        booktabs = TRUE,
        digits = 4,
        caption = "\\label{tab:model-results-log-incidence}Dependent variable is log(Incidence).",
        align = c("l", "c", "c"),
        col.names = c("Variable", "Coefficient Estimate", "p-value")) %>%
  kable_styling(latex_options = c("striped"), font_size = 7) %>%
  footnote(general = c(paste("Coefficient of Determination $R^2$=", round(summary(mod1)$r.squared, 3)),
                       paste("Adjusted Coefficient of Determination $R^2$=", round(summary(mod1)$adj.r.squared, 3)),
                       paste("Standard Error $\\\\sigma$=", round(summary(mod1)$sigma, 3))),
           escape = FALSE)
```

I am also concerned that almost all the variables in the model have a quadratic form and by the number of interactions in the model. Is this an artifact of not controlling for the population size effect?

>\textcolor{blue}{The quadratic form is to capture trends in attribute space. As the performance of the model shows, these terms are effective at this task. And seeing how the terms are significant even after controlling for population size (i.e., modelling the incidence per 100,000 instead of new cases), I hope that your concerns will be allayed.}

>\textcolor{blue}{Finally, I have taken the opportunity of this revision to update the analysis to May 5, 2020.}
